---
author: Daniel Evans
date: January 21, 2018
output: html_document
title: Chapter_-1
---

<h1>Table of Contents<span class="tocSkip"></span></h1>
<div class="toc" style="margin-top: 1em;"><ul class="toc-item"></ul></div>

c func makes array of values

```{r}
x <- c(10, 3, 3, 4)
x + 4
```

```{r}
y <- max(x)
y
```

Math on vectors computes for each element

```{r}
x^2
```

Take the mean of vector

```{r}
x <- c(1, 3, 4, 6, 5, 6)
mean(x)
```

Standard deviation of vector

```{r}
y <- matrix(x, nrow=3)
y
```

Applies sum function to each row in matrix

```{r}
apply(y, 1, sum)
```

Applies sum function to each column in matrix

```{r}
apply(y, 2, sum)
```

```{r}
z <- matrix(x, nrow=3, byrow=TRUE)
z
```

```{r}
name <- c("Emily", "Tim", "Jeff")
# This will combine the z matrix with the name (makes three columns)
my.mat <- cbind(name, z)
my.mat
```

```{r}
is.matrix(my.mat)
```

```{r}
is.data.frame(my.mat)
```

```{r}
my.mat[,2]
```

Even though my.mat[,2] is numbers, since it contains strings as well, numbers are coerced to strings so we can't operate on them as numbers

```{r}
mean(my.mat[,2])
```

Data frames can have different types of data

```{r}
my.data <- data.frame(name, z)
is.data.frame(my.data)
```

So we can take mean as expected

```{r}
mean(my.data[,2])
```

names gets/sets the column names of data frame

```{r}
names(my.data) <- c("Name", "Teeth", "Crawl")
```

```{r}
my.data
```

```{r}
# mean(Teeth) Throws error because Teeth column name not in global scope
```

```{r}
mean(my.data$Teeth)
```

attach() brings all columns in data frame into global scope w/ var names equal to column name

```{r}
attach(my.data)
```

```{r}
mean(Teeth)
```

```{r}
detach(my.data)
```

```{r}
?detach # For when you need help :)
```

```{r}
Data1Survey.df <- read.csv(file='Day1Survey.csv')
```

```{r}
head(Data1Survey.df)
```

```{r}
dim(Data1Survey.df)
```

```{r}
names(Data1Survey.df)
```

```{r}
attach(Data1Survey.df)
```

Get the first row (0 just gets the column names). Of first row get 3, 4 and 6th column. c func === column

```{r}
Data1Survey.df[1, c(1,4,6)]
```

Seems like R indexing starts at 1 :(((((

```{r}
Data1Survey.df[1, c(0,4,6)]
```

```{r}
Data1Survey.df[1, c(1,4,6)]
```

```{r}
head(Data1Survey.df)
```

Get values ALL values (i.e. from every row) in the 3rd column

```{r}
# Just thought about it, was supposed to name this Day1Survey.df, not that it matters...
Data1Survey.df[,3]
```

R vector/dataframe operations work like Python's pandas library :)

```{r}
Sex=="M"
```

Get each row that is a male

```{r}
males.df <- Data1Survey.df[Sex=="M",]
```

```{r}
head(males.df) # Notice how every row for Sex is M, this is true for rest of DataFrame as well 
```

```{r}
dim(males.df)
```

Get only males who have height >= 72, get only the columns "Height" and "TV" -> Seems like R selection statements also work like pandas

```{r}
TallTV.males.df <- Data1Survey.df[Sex=="M" & Height>=72, c("Height", "TV")]
```

```{r}
head(TallTV.males.df)
```

```{r}
exp(1)
```

```{r}
log(exp(4))
```

```{r}
sqrt(20)
```

Functions YAYYYYYY :)

```{r}
mymean <- function(x, trim=0) {
    # Sorts the vector so we can trim outliers
    y <- sort(x) 
    # Add 1 b/c R indexing starts at 1
    # trim is meant to trim outliers 
    # Seems like R vector slicing also works like python
    # But does -1 work to get last element?
    mean(y[(trim+1):(length(y)-trim)])
}
```

```{r}
vec <- c(1,2,3,4,5,6,20)
mymean(vec, trim=1)
```

```{r}
# Seems like this doesn't work like python :(
# Using negative indices removes the element at the positive version of the index
print(vec)
print(vec[-1])
print(vec[-2])
```

```{r}
mymean(vec) # We don't have to pass for default parameter vals
```

```{r}
```